PANCHANGE_analysis

### 2020-08-26 PANCHANGE_analysis.Rmd
cleaning up the disorder chunk

# Age
```{r descriptives Ages}
dat %>%
  group_by(Sample) %>%
  summarise(Mean = mean(Age),
            SD = sd(Age),
            Min = min(Age),
            Max = max(Age))
```

+++ ADD AGE PLOT HERE

+++ADD CHUNK FOR AGE CATEGORIES HERE
```{r Age categories}
dat %>%
  freq(age_category)
```

+++ADD CHUNK FOR AGE CATEGORIES HERE
```{r Age categories by Sample}
dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))
```

```{r Gender per Sample count cross table}
dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n)
```

```{r Gender per Sample count cross table chi square test}
dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-age_category) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0))
    ) %>%
  chisq.test() %>%
  glance()
```

# Psychiatric disorders
```{r vector with psychiatric disorders}
disorder_columns <- c(
    "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "psychotic_disorders_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "autism_spectrum_disorder_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric",
    "control_numeric"
)
```

Correlation matrix of psychiatric disorders
```{r Correlation matrix of psychiatric disorders}
# Select items for correlation matrix
mhd.items <- c(
    "depressive_disorders",
    "anxiety_disorders",
    "eating_disorders",
    "obsessive_compulsive_disorders",
    "psychotic_disorders",
    "mhd.mania_hypomania_bipolar_or_manicdepression",
    "mhd.posttraumatic_stress_disorder_ptsd",
    "autism_spectrum_disorder",
    "mhd.attention_deficit_hyperactivity_disorder",
    "mhd.personality_disorder",
    "control"
)

@@

Define disorder columns in vector

```{r Disorder_inc by Gender}
Disorder_inc_by_Gender <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Gender) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Gender
```

```{r Disorder_inc by Gender prop}
Disorder_inc_by_Gender <- dat.long.psych %>%
  group_by(Disorder_inc) %>%
  filter(Diagnosis == 1) %>%
  count(Gender) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Gender
```

```{r Disorder_inc by Gender plot}
Disorder_inc_by_Gender %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
    ) +
  geom_bar(
    aes(fill = Gender),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Gender") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

@@

Compare the GLAD/EDGI and COPING cohorts as a whole
+++CH: Diagnosis, ethnicity, gender, 
+++MD: May not be possible for NBR; potentially send them code to run it for us

>>>>

# Psychiatric disorders

## Vector
```{r vector with psychiatric disorders}
disorder_columns <- c(
    "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "psychotic_disorders_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "autism_spectrum_disorder_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric",
    "control_numeric"
)
```

## Correlation matrix
```{r Correlation matrix of psychiatric disorders}
# Select items for correlation matrix
mhd.items <- c(
    "depressive_disorders",
    "anxiety_disorders",
    "eating_disorders",
    "obsessive_compulsive_disorders",
    "psychotic_disorders",
    "mhd.mania_hypomania_bipolar_or_manicdepression",
    "mhd.posttraumatic_stress_disorder_ptsd",
    "autism_spectrum_disorder",
    "mhd.attention_deficit_hyperactivity_disorder",
    "mhd.personality_disorder",
    "control"
)

@@

```{r Disorder_inc by Gender_collapsed}
Disorder_inc_by_Gender_collapsed <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Gender_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Gender_collapsed
```

```{r Disorder_inc by Gender_collapsed prop}
Disorder_inc_by_Gender_collapsed <- dat.long.psych %>%
  group_by(Disorder_inc) %>%
  filter(Diagnosis == 1) %>%
  count(Gender_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))

Disorder_inc_by_Gender_collapsed
```

```{r Disorder_inc by Gender_collapsed count plot}
Disorder_inc_by_Gender_collapsed_count_plot <- Disorder_inc_by_Gender_collapsed %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
    ) +
  geom_bar(
    aes(fill = Gender_collapsed),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Gender_collapsed") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4
    ) +
  coord_flip()

Disorder_inc_by_Gender_collapsed_count_plot
```

```{r Disorder_inc by Gender_collapsed prop plot}
Disorder_inc_by_Gender_collapsed_prop_plot <- Disorder_inc_by_Gender_collapsed %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = Prop)
    ) +
  geom_bar(
    aes(fill = Gender_collapsed),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Gender_collapsed") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4
    ) +
  coord_flip()

Disorder_inc_by_Gender_collapsed_prop_plot
```

@@

Compare the GLAD/EDGI and COPING cohorts as a whole
+++CH: Diagnosis, ethnicity, gender, age

# age_category
```{r age_category per Sample prop per total}
age_category_per_Sample_prop_total <- dat %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(age_category_per_Sample_prop_total)
age_category_per_Sample_prop_total
```

```{r age_category per Sample prop per Sample}
age_category_per_Sample_prop_Sample <- dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2))

kable(age_category_per_Sample_prop_Sample)
age_category_per_Sample_prop_Sample
```

```{r age_category per Sample prop cross table}
age_category_per_Sample_prop_Sample_cross_table <- dat %>%
  group_by(Sample) %>%
  count(age_category, Sample) %>%
  mutate(Prop = round(n/sum(n), 2)) %>%
  pivot_wider(names_from = Sample, values_from = c(n, Prop))

kable(age_category_per_Sample_prop_Sample_cross_table)
age_category_per_Sample_prop_Sample_cross_table
```

```{r age_category per Sample count cross table}
age_category_per_Sample_count_cross_table <- dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  )

kable(age_category_per_Sample_count_cross_table)
age_category_per_Sample_count_cross_table
```

```{r age_category per Sample chisquare test}
age_category_per_Sample_chissquare_test <- dat %>%
  count(age_category, Sample) %>%
  spread(key = Sample, value = n) %>%
  select(-age_category) %>%
  mutate(
    EDGI = as.integer(replace_na(EDGI, 0)),
    GLAD = as.integer(replace_na(GLAD, 0)),
    NBR = as.integer(replace_na(NBR, 0)),
    RAMP = as.integer(replace_na(RAMP, 0))
  ) %>%
  chisq.test() %>%
  glance()

kable(age_category_per_Sample_chissquare_test)
age_category_per_Sample_chissquare_test
```

```{r age_category by Sample count plot}
age_category_by_Sample_count_plot <- age_category_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = age_category,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "age_category",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

age_category_by_Sample_count_plot
```

```{r age_category by Sample prop plot}
age_category_by_Sample_prop_plot <- age_category_per_Sample_prop_Sample %>%
  ggplot(mapping = aes(
    x = age_category,
    y = Prop)
  ) +
  geom_bar(
    aes(fill = Sample),
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "age_category",
    y = "Percentage per sample",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6
  ) +
  coord_flip()

age_category_by_Sample_prop_plot
```
